---
title: "Semi Bernoulli Items"
author: "Ben Ewing and Neil Pruthi"
date: "December 7, 2018"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

# Setup

R setup.

```{r r-setup, include = FALSE}
# For data manipulation
library(dplyr)
library(purrr)
library(tidyr)
library(e1071)
# For linear programming
library(ROI)
library(ROI.plugin.glpk)
library(ompr)
library(ompr.roi)
# Plotting
library(ggplot2)
library(ggthemes)
theme_set(theme_tufte())
# Minimax
source("../r/n_player_minimax.R")
# Simulation settings
n_draws <- 1000
```

Python/reticulate setup.

```{r py-setup}
library(reticulate)
use_python("C:/ProgamData/Anaconda3/python.exe")
# Load utility functions
source_python("../py/estimate_utility.py")
# and Numpy
np <- import("numpy", convert=FALSE)
```

```{r three-player}
n_players <- as.integer(3)
```

## Normal Items

Will minimax attempt to approximate a uniform distribution when given a set of normal items, with three players?

### Expected Value: 0

It is equally likely for an item to take positive and negative values.

```{r exp0-density, warning = FALSE, cache = TRUE}
# Generate items
items <- sapply(seq(2, 10, 0.15), function(x) rnorm(n_draws, 0, x))

# A density plot
items %>% 
  as_tibble() %>% 
  gather(item, val) %>% 
  ggplot(aes(val, colour = item, fill = item)) +
  geom_density(alpha = 0.10) +
  xlim(c(-35, 35)) +
  theme(legend.position = "none", axis.title = element_blank())

# Estimate utility
items_u <- est_knap_u(items, n_players)

# And run minimax
mini <- minimax(items_u, n_players)
mini$solution

# Minimax components
mini_comp <- paste0("V", which(mini$solution[-length(mini$solution)] > 0))

# Create a mixture distribution and append to the end of items
items_aug <- cbind(
  items,
  map(1:(length(mini$solution) - 1), ~ {
    if (mini$solution[.x] > 0)
      items[,.x][1:ceiling(n_draws*mini$solution[.x])]
    else
      numeric()
  }) %>% 
    unlist() %>% 
    .[1:n_draws]
)

# To dataframe
items_aug <- items_aug %>% 
  as_data_frame() %>% 
  gather(item, val) %>% 
  mutate(item = ifelse(item == paste0("V", (ncol(items) + 1)), "Minimax", item),
         mark = case_when(item == "Minimax" ~ "Minimax",
                          item %in% mini_comp ~ "Minimax Component",
                          T ~ "Not Used by Minimax"))

# Plot again
ggplot(items_aug, aes(val, fill = item, colour = item)) +
  geom_density(alpha = 0.3) +
  facet_wrap(. ~ mark, ncol = 1) +
  theme(legend.position = "none")

# Summary statistics
items_aug %>% 
  group_by(item) %>% 
  summarise(mean = mean(val), sd = sd(val), skewness = skewness(val), kurtosis = kurtosis(val),
            mark = unique(mark)) %>% 
  arrange(mark)
```

### Expected Value: 10

Items will mostly take on positive values.

```{r exp10-density, warning = FALSE, cache = TRUE}
# Generate items
items <- sapply(seq(2, 10, 0.15), function(x) rnorm(n_draws, 10, x))

# A density plot
items %>% 
  as_tibble() %>% 
  gather(item, val) %>% 
  ggplot(aes(val, colour = item, fill = item)) +
  geom_density(alpha = 0.10) +
  xlim(c(-35, 35)) +
  theme(legend.position = "none", axis.title = element_blank())

# Estimate utility
items_u <- est_knap_u(items, n_players)

# And run minimax
mini <- minimax(items_u, n_players)
mini$solution

# Minimax components
mini_comp <- paste0("V", which(mini$solution[-length(mini$solution)] > 0))

# Create a mixture distribution and append to the end of items
items_aug <- cbind(
  items,
  map(1:(length(mini$solution) - 1), ~ {
    if (mini$solution[.x] > 0)
      items[,.x][1:ceiling(n_draws*mini$solution[.x])]
    else
      numeric()
  }) %>% 
    unlist() %>% 
    .[1:n_draws]
)

# To dataframe
items_aug <- items_aug %>% 
  as_data_frame() %>% 
  gather(item, val) %>% 
  mutate(item = ifelse(item == paste0("V", (ncol(items) + 1)), "Minimax", item),
         mark = case_when(item == "Minimax" ~ "Minimax",
                          item %in% mini_comp ~ "Minimax Component",
                          T ~ "Not Used by Minimax"))

# Plot again
ggplot(items_aug, aes(val, fill = item, colour = item)) +
  geom_density(alpha = 0.3) +
  facet_wrap(. ~ mark, ncol = 1) +
  theme(legend.position = "none")

# Summary statistics
items_aug %>% 
  group_by(item) %>% 
  summarise(mean = mean(val), sd = sd(val), skewness = skewness(val), kurtosis = kurtosis(val),
            mark = unique(mark)) %>% 
  arrange(mark)
```

## Uniform Items

### Same mean.

```{r uniform-items, warning = FALSE, cache = TRUE}
# Generate items
items <- sapply(seq(2, 10, 0.5), function(x) runif(n_draws, -x, x))

# A density plot
items %>% 
  as_tibble() %>% 
  gather(item, val) %>% 
  ggplot(aes(val, colour = item, fill = item)) +
  geom_density(alpha = 0.10) +
  xlim(c(-35, 35)) +
  theme(legend.position = "none", axis.title = element_blank())

# Estimate utility
items_u <- est_knap_u(items, n_players)

# And run minimax
mini <- minimax(items_u, n_players)
mini$solution

# Minimax components
mini_comp <- paste0("V", which(mini$solution[-length(mini$solution)] > 0))

# Create a mixture distribution and append to the end of items
items_aug <- cbind(
  items,
  map(1:(length(mini$solution) - 1), ~ {
    if (mini$solution[.x] > 0)
      items[,.x][1:ceiling(n_draws*mini$solution[.x])]
    else
      numeric()
  }) %>% 
    unlist() %>% 
    .[1:n_draws]
)

# To dataframe
items_aug <- items_aug %>% 
  as_data_frame() %>% 
  gather(item, val) %>% 
  mutate(item = ifelse(item == paste0("V", (ncol(items) + 1)), "Minimax", item),
         mark = case_when(item == "Minimax" ~ "Minimax",
                          item %in% mini_comp ~ "Minimax Component",
                          T ~ "Not Used by Minimax"))

# Plot again
ggplot(items_aug, aes(val, fill = item, colour = item)) +
  geom_density(alpha = 0.3) +
  facet_wrap(. ~ mark, ncol = 1) +
  theme(legend.position = "none")

# Summary statistics
items_aug %>% 
  group_by(item) %>% 
  summarise(mean = mean(val), sd = sd(val), skewness = skewness(val), kurtosis = kurtosis(val),
            mark = unique(mark)) %>% 
  arrange(mark)
```

What happens with more players?

```{r uniform-items-more-players, warning = FALSE, cache = TRUE}
n_players <- as.integer(5)

# Generate items
items <- sapply(seq(2, 10, 0.5), function(x) runif(n_draws, -x, x))

# A density plot
items %>% 
  as_tibble() %>% 
  gather(item, val) %>% 
  ggplot(aes(val, colour = item, fill = item)) +
  geom_density(alpha = 0.10) +
  xlim(c(-35, 35)) +
  theme(legend.position = "none", axis.title = element_blank())

# Estimate utility
items_u <- est_knap_u(items, n_players)

# And run minimax
mini <- minimax(items_u, n_players)
mini$solution

# Minimax components
mini_comp <- paste0("V", which(mini$solution[-length(mini$solution)] > 0))

# Create a mixture distribution and append to the end of items
items_aug <- cbind(
  items,
  map(1:(length(mini$solution) - 1), ~ {
    if (mini$solution[.x] > 0)
      items[,.x][1:ceiling(n_draws*mini$solution[.x])]
    else
      numeric()
  }) %>% 
    unlist() %>% 
    .[1:n_draws]
)

# To dataframe
items_aug <- items_aug %>% 
  as_data_frame() %>% 
  gather(item, val) %>% 
  mutate(item = ifelse(item == paste0("V", (ncol(items) + 1)), "Minimax", item),
         mark = case_when(item == "Minimax" ~ "Minimax",
                          item %in% mini_comp ~ "Minimax Component",
                          T ~ "Not Used by Minimax"))

# Plot again
ggplot(items_aug, aes(val, fill = item, colour = item)) +
  geom_density(alpha = 0.3) +
  facet_wrap(. ~ mark, ncol = 1) +
  theme(legend.position = "none")

# Summary statistics
items_aug %>% 
  group_by(item) %>% 
  summarise(mean = mean(val), sd = sd(val), skewness = skewness(val), kurtosis = kurtosis(val),
            mark = unique(mark)) %>% 
  arrange(mark)
```

Items don't overlap (three players again).

```{r uniform-items-2, warning = FALSE, cache = TRUE}
n_players <- as.integer(3)

# Generate items
items <- sapply(seq(2, 10, 0.5), function(x) runif(n_draws, x, x+0.499))

# A density plot
items %>% 
  as_tibble() %>% 
  gather(item, val) %>% 
  ggplot(aes(val, colour = item, fill = item)) +
  geom_density(alpha = 0.10) +
  # xlim(c(-35, 35)) +
  theme(legend.position = "none", axis.title = element_blank())

# Estimate utility
items_u <- est_knap_u(items, n_players)

# And run minimax
mini <- minimax(items_u, n_players)
mini$solution

# Minimax components
mini_comp <- paste0("V", which(mini$solution[-length(mini$solution)] > 0))

# Create a mixture distribution and append to the end of items
items_aug <- cbind(
  items,
  map(1:(length(mini$solution) - 1), ~ {
    if (mini$solution[.x] > 0)
      items[,.x][1:ceiling(n_draws*mini$solution[.x])]
    else
      numeric()
  }) %>% 
    unlist() %>% 
    .[1:n_draws]
)

# To dataframe
items_aug <- items_aug %>% 
  as_data_frame() %>% 
  gather(item, val) %>% 
  mutate(item = ifelse(item == paste0("V", (ncol(items) + 1)), "Minimax", item),
         mark = case_when(item == "Minimax" ~ "Minimax",
                          item %in% mini_comp ~ "Minimax Component",
                          T ~ "Not Used by Minimax"))

# Plot again
ggplot(items_aug, aes(val, fill = item, colour = item)) +
  geom_density(alpha = 0.3) +
  facet_wrap(. ~ mark, ncol = 1) +
  theme(legend.position = "none")

# Summary statistics
items_aug %>% 
  group_by(item) %>% 
  summarise(mean = mean(val), sd = sd(val), skewness = skewness(val), kurtosis = kurtosis(val),
            mark = unique(mark)) %>% 
  arrange(mark)
```

## Mix of items

All have random means (within bounds),

```{r dist-funcs}
dist_functions <- list(
  wrnorm = function(n) rnorm(n, runif(1, 0, 5), runif(1, 0, 5)),
  wrunif = function(n) {
    min <- runif(1, -5, 5)
    max <- min + runif(1, -5, 5)
    max <- ifelse(max < min, abs(max)+min, max)
    runif(n, min, max)
  },
  wrgamma = function(n) rgamma(n, runif(1, 0, 5))
)
```

```{r mix-items}
# Generate items
items <- sapply(seq(2, 10, 0.5), function(x) runif(n_draws, x, x+0.499))

# A density plot
items %>% 
  as_tibble() %>% 
  gather(item, val) %>% 
  ggplot(aes(val, colour = item, fill = item)) +
  geom_density(alpha = 0.10) +
  # xlim(c(-35, 35)) +
  theme(legend.position = "none", axis.title = element_blank())

# Estimate utility
items_u <- est_knap_u(items, n_players)

# And run minimax
mini <- minimax(items_u, n_players)
mini$solution

# Minimax components
mini_comp <- paste0("V", which(mini$solution[-length(mini$solution)] > 0))

# Create a mixture distribution and append to the end of items
items_aug <- cbind(
  items,
  map(1:(length(mini$solution) - 1), ~ {
    if (mini$solution[.x] > 0)
      items[,.x][1:ceiling(n_draws*mini$solution[.x])]
    else
      numeric()
  }) %>% 
    unlist() %>% 
    .[1:n_draws]
)

# To dataframe
items_aug <- items_aug %>% 
  as_data_frame() %>% 
  gather(item, val) %>% 
  mutate(item = ifelse(item == paste0("V", (ncol(items) + 1)), "Minimax", item),
         mark = case_when(item == "Minimax" ~ "Minimax",
                          item %in% mini_comp ~ "Minimax Component",
                          T ~ "Not Used by Minimax"))

# Plot again
ggplot(items_aug, aes(val, fill = item, colour = item)) +
  geom_density(alpha = 0.3) +
  facet_wrap(. ~ mark, ncol = 1) +
  theme(legend.position = "none")

# Summary statistics
items_aug %>% 
  group_by(item) %>% 
  summarise(mean = mean(val), sd = sd(val), skewness = skewness(val), kurtosis = kurtosis(val),
            mark = unique(mark)) %>% 
  arrange(mark)
```

# Player to item ratio

```{r uniform-items-2, warning = FALSE, cache = TRUE}
n_players <- as.integer(3)
n_draws <- 1000
n_items <- 10

# Generate items
# items <- sapply(seq(1, 10, length.out = n_items), function(x) runif(n_draws, -x, x))
# items <- np_array(sapply(seq(1, 10, length.out = n_items), function(x) sample(seq(-x, x, length.out = n_draws))))

# Try to sim the mean var tradeoff
items <- np_array(cbind(
  sample(seq(-1, 1, length.out = n_draws)),
  sample(seq(-2, 2, length.out = n_draws)),
  sample(seq(-13.35, 13.35, length.out = n_draws))
))

# A density plot
# items %>% 
#   as.matrix() %>% 
#   as_tibble() %>% 
#   gather(item, val) %>% 
#   ggplot(aes(val, colour = item, fill = item)) +
#   geom_density(alpha = 0.10) +
#   theme(legend.position = "none", axis.title = element_blank())

# Estimate utility
items_u <- est_knap_u(items, n_players)

# And run minimax
mini <- minimax(list2env(items_u), n_players)
mini$solution

# Minimax components
mini_comp <- paste0("V", which(mini$solution[-length(mini$solution)] > 0))

# minimax probabilities
mini_pr <- data_frame(
  pr = mini$solution[-length(mini$solution)],
  item = paste0("V", 1:(length(mini$solution)-1))
)

items <- as.matrix(items)
# Create a mixture distribution and append to the end of items
items_aug <- cbind(
  items,
  map(1:(length(mini$solution) - 1), ~ {
    if (mini$solution[.x] > 0)
      items[,.x][1:ceiling(n_draws*mini$solution[.x])]
    else
      numeric()
  }) %>% 
    unlist() %>% 
    .[1:n_draws]
)

# To dataframe
items_aug <- items_aug %>% 
  as_data_frame() %>% 
  gather(item, val) %>% 
  mutate(item = ifelse(item == paste0("V", (ncol(items) + 1)), "Minimax", item),
         mark = case_when(item == "Minimax" ~ "Minimax",
                          item %in% mini_comp ~ "Minimax Component",
                          T ~ "Not Used by Minimax"))

# Plot again
ggplot(items_aug, aes(val, fill = item, colour = item)) +
  geom_density(alpha = 0.3) +
  facet_wrap(. ~ mark, ncol = 1) +
  theme(legend.position = "none")

# Summary statistics
items_aug_sum <- items_aug %>% 
  group_by(item) %>% 
  summarise(mean = mean(val), sd = sd(val), skewness = skewness(val), kurtosis = kurtosis(val),
            mark = unique(mark)) %>% 
  left_join(mini_pr, by = "item") %>% 
  arrange(mark)
items_aug_sum

# Look at qualities 
items_aug_sum %>% 
  select(-item, -mark) %>% 
  gather(measure, val, -pr) %>% 
  ggplot(aes(pr, val)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(. ~ measure, scales = "free", ncol = 2)
```

Mean variance tradeoff.

```{r mean-var-test}
n_draws <- 1000
map_dfr(rep(seq(13, 13.5, 0.1), each = 10), ~ {
  np_array(cbind(
    runif(n_draws, -1, 1), # sample(seq(-1, 1, length.out = n_draws)),
    runif(n_draws, -2, 2), # sample(seq(-2, 2, length.out = n_draws)),
    runif(n_draws, -.x, .x) # sample(seq(-.x, .x, length.out = n_draws))
  )) %>% 
    est_knap_u(as.integer(3)) %>% 
    minimax(3) %>% 
    .$solution %>% 
    as_data_frame() %>% 
    tibble::rownames_to_column() %>% 
    spread(rowname, value) %>% 
    mutate(a = .x)
}) -> tmp

tmp %>% 
  mutate(a = round(a, digits = 1)) %>%
  group_by(a) %>% 
  summarise_if(is.numeric, mean)
```