---
title: "Player to Knapsack Ratio"
author: "Ben Ewing"
date: "November 25, 2018"
output:
  pdf_document: default
header-includes:
  - \usepackage{amsfonts}
  - \usepackage{stmaryrd}
  - \usepackage{bm}
  - \usepackage{bbm}
geometry: margin=2cm
editor_options: 
  chunk_output_type: console
---

# Setup

R.

```{r r-setup, include = F}
library(dplyr)
library(purrr)
library(tidyr)
library(RcppAlgos)
library(ggplot2)
```


Helper functions.
```{r heleprs}
n_ties <- function(n_players, n_knaps) {
  data_frame(
    n_players = n_players,
    n_knaps = n_knaps,
    player_knap_ratio = n_players/n_knaps,
    n_matches = n_knaps^n_players,
    n_complete_ties = n_knaps,
    n_any_ties = n_matches - (factorial(n_knaps)/factorial(n_knaps-n_players)),
    n_any_ties_n_matches_ratio = n_any_ties/n_matches
    # tie_matchup_ratio = n_ties/n_matches
  )
}
```

# Discussion

Recall that if a player ties and wins (a split win), they split their winnings with all tied players. For example, in a game where two players win and one player loses, the winning players each get a utility of 0.5, and the loser -1. Thus, a win is strictly better than a split win, which is strictly better than a loss. 

Observe that if there are $n$ players $m$ items, there are $m^n$ possible matches, of which there are:
  * $m$ matches where all players play the same item
    * I can't really figure out the expansion here
    * i.e. how many matches where $n-1$ players all play the same item (but not the same item as the first player)?
  * $\frac{m!}{(m-n)!$ matches where all players play a unique item

The maximum achievable utility for the tied players (i.e. in the case where they win) is $0, \frac{1}{n-1}, \frac{2}{n-2}, \ldots, \frac{n}{n-(n-1)}$ (in the first case, all tie, in the last case, only one person wins). 


