---
title: "Simultaneous Adversarial Knapsack"
author: "Ben Ewing"
date: "November 25, 2018"
output:
  pdf_document: default
header-includes:
  - \usepackage{amsfonts}
  - \usepackage{stmaryrd}
  - \usepackage{bm}
  - \usepackage{bbm}
geometry: margin=2cm
editor_options: 
  chunk_output_type: console
---

# Setup

Libraries.

```{r libraries, message = F}
# For data manipulation
library(dplyr)
library(purrr)
library(tidyr)
# For kurtosis and skew functions
library(e1071)
# For fast combinations
library(RcppAlgos)
# For compiling functions to bytecode, this usually nets a tiny speedup
library(compiler)
# For linear programming
library(ROI)
library(ROI.plugin.glpk)
library(ompr)
library(ompr.roi)
# For plots and tables
library(ggplot2)
library(knitr)
library(kableExtra)
# For parallelization
library(parallel)
```

Load knapsack functions.

```{r knapsack-funcs}
source("../r/knapsack_functions_mat.R")
```

Simulation settings.

```{r sim-settings}
n_knap_sims <- 500
n_draws <- 1000
n_items <- 20
knap_size <- 1
```

The set of distributions an item can draw from.
```{r dist-functions}
dist_functions <- list(
  wrnorm = function(n) rnorm(n, runif(1, 0, 5), runif(1, 0, 5)),
  wrunif = function(n) {
    min <- runif(1, -5, 5)
    max <- min + runif(1, -5, 5)
    max <- ifelse(max < min, abs(max)+min, max)
    runif(n, min, max)
  },
  wrgamma = function(n) rgamma(n, runif(1, 0, 5))
)
```

# Goal

To understand why minimax chooses the strategies that it does.

# Approach

For each scenario, simulate many games. Within each game find the minimax strategy. Then look at descripives of minimax winners.

# Two Players

Set two players.

```{r two-player-settings}
n_players <- 2
```

Run simulations.

```{r two-player-knaps}
if (!file.exists("../sims/two_player_minimax_results.Rds")) {
  # In parallel?
  cl <- makeCluster(16)
  clusterExport(cl, c("n_items", "n_players", "dist_functions", "n_draws", "estimate_knapsack_utility",
                      "strat_minimax_2"))
  clusterEvalQ(cl, {
    # For data manipulation
    library(dplyr)
    library(purrr)
    library(tidyr)
    # For kurtosis and skew functions
    library(e1071)
    # For fast combinations
    library(RcppAlgos)
    # For compiling functions to bytecode, this usually nets a tiny speedup
    library(compiler)
    # For linear programming
    library(ROI)
    library(ROI.plugin.glpk)
    library(ompr)
    library(ompr.roi)
  })
  
  knaps <- parLapply(cl, 1:n_knap_sims, function(.x) {
    mat <- replicate(n_items, dist_functions[[sample(1:length(dist_functions), 1)]](n_draws))
    minimax <- estimate_knapsack_utility(mat, n_players) %>%
      strat_minimax_2() %>%
      group_by(knap_id) %>%
      summarise_if(is.numeric, mean) %>%
      arrange(knap_id) %>% mutate(knap_id = as.character(knap_id))
    mat_minimax_weighted <- apply(mat, 1, function(x) weighted.mean(x, minimax$probability))
    as_data_frame(mat) %>%
      mutate(minimax_weighted = mat_minimax_weighted) %>%
      gather(knap_id, value) %>%
      mutate(knap_id = gsub("V", "", knap_id)) %>%
      left_join(select(minimax, knap_id, utility, probability), by = "knap_id") %>%
      mutate(knap_id = paste0(knap_id, "_", .x), sim_id = .x)
  })
  
  stopCluster(cl)
  
  knaps <- bind_rows(knaps) %>% 
    group_by(knap_id) %>%
    mutate(
      mean = mean(value), sd = sd(value),
      skewness = skewness(value), kurtosis = kurtosis(value)
    )
  saveRDS(knaps, "../sims/two_player_minimax_results.Rds")
} else {
  knaps <- readRDS("../sims/two_player_minimax_results.Rds")
}
```

Plot knapsack descriptives for knapsacks that received positive probability.

```{r 2p-knapsack-stats}
knaps %>% 
  # this also drops minimax_weighted rows, they have NA probability
  filter(probability > 0) %>% 
  summarise(probability = unique(probability), mean = unique(mean), sd = unique(sd), 
            skewness = unique(skewness), kurtosis = unique(kurtosis)) %>% 
  gather(measure, val, -knap_id, -probability) %>% 
  ggplot(aes(probability, val)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  facet_wrap(. ~ measure, scales = "free")
```

<!-- Plot distributions? Pretty messy. -->

```{r knapsack-dists}
# knaps %>% 
#   select(sim_id, knap_id, value, probability) %>% 
#   mutate(label = case_when(
#     probability > 0 ~ "minimax_part",
#     grepl("minimax", knap_id) ~ "minimax",
#     T ~ "Other"
#   )) %>% 
#   ggplot(aes(value, group = knap_id, fill = knap_id, colour = knap_id)) +
#   geom_density(alpha = 0.3, show.legend = F) +
#   facet_grid(sim_id ~ label, scales = "free")
```

# Three Players

Set three players.

```{r three-player-settings}
n_players <- 3
```

Run simulations.

```{r three-player-knaps}
if (!file.exists("../sims/three_player_minimax_results.Rdata")) {
  # In parallel?
  cl <- makeCluster(16)
  clusterExport(cl, c("n_items", "n_players", "dist_functions", "n_draws", "estimate_knapsack_utility",
                      "strat_minimax_3"))
  clusterEvalQ(cl, {
    # For data manipulation
    library(dplyr)
    library(purrr)
    library(tidyr)
    # For kurtosis and skew functions
    library(e1071)
    # For fast combinations
    library(RcppAlgos)
    # For compiling functions to bytecode, this usually nets a tiny speedup
    library(compiler)
    # For linear programming
    library(ROI)
    library(ROI.plugin.glpk)
    library(ompr)
    library(ompr.roi)
  })
  
  knaps <- parLapply(cl, 1:n_knap_sims, function(.x) {
    mat <- replicate(n_items, dist_functions[[sample(1:length(dist_functions), 1)]](n_draws))
    minimax <- estimate_knapsack_utility(mat, n_players) %>%
      strat_minimax_3() %>%
      group_by(knap_id) %>%
      summarise_if(is.numeric, mean) %>%
      arrange(knap_id) %>% mutate(knap_id = as.character(knap_id))
    mat_minimax_weighted <- apply(mat, 1, function(x) weighted.mean(x, minimax$probability))
    as_data_frame(mat) %>%
      mutate(minimax_weighted = mat_minimax_weighted) %>%
      gather(knap_id, value) %>%
      mutate(knap_id = gsub("V", "", knap_id)) %>%
      left_join(select(minimax, knap_id, utility, probability), by = "knap_id") %>%
      mutate(knap_id = paste0(knap_id, "_", .x), sim_id = .x)
  })
  
  stopCluster(cl)
  
  knaps <- bind_rows(knaps) %>% 
    group_by(knap_id) %>%
    mutate(
      mean = mean(value), sd = sd(value),
      skewness = skewness(value), kurtosis = kurtosis(value)
    )
  saveRDS(knaps, "../sims/three_player_minimax_results.Rds")
} else {
  knaps <- readRDS("../sims/three_player_minimax_results.Rds")
}
```

Plot knapsack descriptives for knapsacks that received positive probability.

```{r 3p-knapsack-stats}
knaps %>% 
  # this also drops minimax_weighted rows, they have NA probability
  filter(probability > 0) %>% 
  summarise(probability = unique(probability), mean = unique(mean), sd = unique(sd), 
            skewness = unique(skewness), kurtosis = unique(kurtosis)) %>% 
  gather(measure, val, -knap_id, -probability) %>% 
  ggplot(aes(probability, val)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  facet_wrap(. ~ measure, scales = "free")
```
